# 4 Analysis


```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```

## 4.1 Research question 1

- Let's put ourselves in context. Our first research question is as follows:

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

RQ

</div>

<br><br><br>

...

<br>


## Unsupervised Learning

Start the unsupervised learning by giving an overlook to the correlation between variables. it is important to look at the correlation structure in order to take decisions about dimension reductions.

```{r}
library(ggcorrplot)

ggcorrplot(cor(final_df[1:16]))


```
We can notice that there are some variable highly negative correlated between them, such as rate_negative_words and rate_positive_words and n_unique_token together with n_token_content. Rather positively correlated are rate_positive_words and global_sentiment polarity. Color gradation gives us the information about the positive/negative correlation intensity.


### Principal Component Analysis (PCA)

After inspecting the correlation structure, we compute the Principal component analysis.
in order to give the same importance to each variable we scale the data first, and then compute the PCA.

```{r}

?PCA
summary(prcomp(final_df[1:16], scale = TRUE))
pop.pca <- PCA(final_df[1:16], ncp = 21, graph = FALSE)
fviz_pca_var(pop.pca)
fviz_eig(pop.pca, addlabels = TRUE, ncp=21)
for (i in 1:6) {
  print(fviz_contrib(pop.pca, choice = "var", axes = i))
}
```
Notice that the first dimention account for 18.8% of the total variance, and the second dimention for 17.6%. Since the third component the explained variance is halved with respect to the second component.

* first component:
Mainly positively correlated with n_non_stop_unique_token and n_unique_token and slightly positively correlated with rate_positive_words; negatively correlated with n_token_content, slightly negatively correlated with rate_negative_words.

* second component:
Mainly positively correlated with rate_positive_words and global_sentiment_polarity; negatively correlated with rate_negative_words. Slightly negatively correlated also with n_non_stop_unique_token and n_unique_token.


## Supervised Learning

In this section, we fit different models in order to predict the categorical variable: share_label.

We conducted two supervised learnings: one for binary class and one for multi class. In the binary class analysis we predict the level of popularity ( given by the variable share_label) that can be: Popular or unpopular.
Whereas, in the multi-classes we predict the share_class variable that can assume one more labels of popularity which is: Outstanding. So in total: Unpopular, popular and outstanding.

### Binary Classes

As always, we split the dataset in 2 subgroups: the training set and the test set. We decided to made a training set that accounts for the 80% of the total dataset, thereby the test set is the 20% remaining.

```{r}
set.seed(1996) ## for replication purpose
#split the data set in training/test set
index.tr <-
  createDataPartition(y = final_df$share_label,
                      p = 0.8, #trainset = 80% of the total dataset
                      list = FALSE)
final_df.tr <- final_df[index.tr,]
final_df.te <- final_df[-index.tr,]
```

With the aim of avoid overfitting, we will use also a cross validation splitting. To be precise, a 5 fold Cross-validation in order to reduce computational time due to our big dataset.

```{r}
trControl <- trainControl(method  = "cv",
                          number  = 5) #5 fold Cross-Validation
```

#### K-nearest neighbor (KNN)

The first model that we are going to present is the KNN.
We predict the categorical variable share_label with all the features in our dataset. It is important to decide which value of the hyperparameter K the model should use in order to optimize our results, to do this we compute a tuning of K taking into account a defined rage of values that K could assume: from 1 to 10. 

From the hyperparameter tuning results, it turns out that the value of K = 9 maximizes Kappa and the accuracy.

```{r}
###KNN 
fit <- train(share_label ~ .,
             #trying to predict share_label with all other features
             method     = "knn",
             tuneGrid   = expand.grid(k = 1:10), #tuning K hyperparameters
             trControl  = trControl,
             metric     = "Accuracy",
             data       = final_df)
#k=9 maximizes our Kappa & Accuracy
final_df.knn <- knn3(data = final_df.tr,
                     share_label ~.,
                     k = 9)
final_df.knn.pred <- predict(final_df.knn,
                             newdata = final_df.te,
                             type = "class")
confusionMatrix(data = as.factor(final_df.knn.pred),
                reference = final_df.te$share_label)
```

As we can see from the results, the prediction quality of the KNN model is not satisfying, we achieve an accuracy of 0.558, even the Kappa has a low value.
Notice that the data seems balanced, indeed sensitivity and specificity take on similar values. by the way, as expected the "lazy learner" provide a really poor prediction.



#### Naive Bayes

Computation of naive Bayes to predict share_label variable using kernel distribution of data.
As always, we train the model on the training set and compute the prediction on the test set.

```{r}
#NAIVE BAYES without hyperparameter tuning
#since it only achieve a 0.177, we think that other algorithms can do better
final_df.nb <- naive_bayes(share_label ~ .,
                           data = final_df.tr,
                           kernel = TRUE)
final_df.nb.pred <- predict(final_df.nb,
                            newdata = final_df.te)
confusionMatrix(data = as.factor(final_df.nb.pred),
                reference = final_df.te$share_label)
```

From the confusion matrix output we notice an accuracy of 0.599; so a slight improvement in the prediction quality compared to KNN.


#### Support Vector Machine (SVM)

As our RAM/CPU struggles while hyper-parameter tuning, we've chosen to sample a part of our data in order to have less computational effort. We fit the SVM model in this sampled dataset.

```{r}
#as our RAM/CPU struggles while hyper-parameter tuning, we've chosen to sample a part of our data in order to have less computational effort
sub_hyp <- sample(nrow(final_df), 5000)
sub_hyp <- final_df[sub_hyp,]
```

We do not assume linear relationship between variables so we compute svm with radial kernel.

Thanks to the tuning of hyperparameters, the values of sigma and C that we will use in the model are chosen: sigma = 0.02 and C = 0.75

```{r}
#SVM
hyperparameter_svm <- expand.grid(sigma = c(0.01, 0.02, 0.05, 0.1),
                           C = seq(0,1,0.25))
svm_hyperpara_train <- train(share_label ~ .,
                         data = sub_hyp,
                         method = "svmRadial",
                         #radial instead of linear since
                         #we assume no linear relationship
                         trControl = trControl,
                         tuneGrid = hyperparameter_svm)
svm_hyperpara_train

#The final values used for the model were sigma = 0.02 and C = 0.75.
final_df.svm_radial <- #train our radial kernel with sigma = 0.02 and C = 0.75
  svm(share_label ~ .,
      data = final_df.tr,
      kernel = "radial",
      tuneGrid = expand.grid(C = 0.75,
                             sigma = 0.02))
final_df.svm.pred_radial <- predict(final_df.svm_radial,
                                    newdata = final_df.te)
confusionMatrix(data = as.factor(final_df.svm.pred_radial),
                reference = final_df.te$share_label)
```
Svm model with radial Kernel, C = 0.75 and sigma = 0.02 gives us an accuracy of 0.63 and Kappa of 0.26.
Although the prediction quality is not totally reliable, the outcome has improved compared to KNN.



#### Classification and regression tree (CART)

Running CART for classification task we find an accuracy of 0.616.

```{r}
final_df.tree <- rpart(share_label ~ .,
                       data = final_df.tr)
rpart.plot(final_df.tree)
final_df.tree.pred <- predict(final_df.tree,
                              newdata = final_df.te,
                              type = "class")
confusionMatrix(data = as.factor(final_df.tree.pred),
                reference = final_df.te$share_label)
#Trying if pruning is needed
plotcp(final_df.tree)
#No need to prune
```
There is no need to prune the tree since we have only 3 nodes and looking at the plot made with plotcp code, we would prune from 3 node to 2; but the model does not improve. So we keep the unpruned tree.


#### Random Forest

```{r}
# HYPERPARAMETER TUNING FOR RF
# THIS CODE WORK PERFECTLY BUT TAKES 20 MIN TO RUN
# WE GIVE YOU THE OUTPUT OF THIS RUN
# IN ORDER TO DECREASE KNIT TIME, WE REMOVE THE CODE AND CONSIDER IT AS TEST
#hyperparameter_rf <-
#  train(
#    share_label ~ .,
#    data = sub_hyp,
#    method = "rf",
#    metric = "Accuracy",
#    tuneGrid = expand.grid(.mtry = seq(1,5, 0.5)),
#    trControl = trControl
#  )
#Random Forest 
#10000 samples
#   21 predictor
#    2 classes: 'Unpopular', 'Popular' 
#No pre-processing
#Resampling: Cross-Validated (5 fold) 
#Summary of sample sizes: 8000, 7999, 8001, 8000, 8000 
#Resampling results across tuning parameters:
#  mtry  Accuracy  Kappa
#  1.0   0.626     0.247
#  1.5   0.639     0.275
#  2.0   0.639     0.277
#  2.5   0.641     0.280
#  3.0   0.633     0.265
#  3.5   0.636     0.271
#  4.0   0.633     0.266
#  4.5   0.634     0.266
#  5.0   0.634     0.268
#Accuracy was used to select the optimal model using the largest value.
#The final value used for the model was mtry = 2.5.
```

The random forest model gives us results really similar to the svm model.
accuracy = 0.637 and Kappa = 0.672
```{r}
rf <- train(share_label~.,
            data = final_df.tr,
            method = "rf",
            metric = "Accuracy",
            tuneGrid = expand.grid(.mtry = 2.5),
            #use the hyperparameter that we've find in
            #the hyperparameter tuning (chunk code line 153-195)
            trControl = trControl)
rf.te.pred <- predict(rf,
                      newdata = final_df.te,
                      type = "raw")
confusionMatrix(data = as.factor(rf.te.pred),
                reference = final_df.te$share_label)
# Best Kappa so far: 0.274
# Best Balanced Accuracy so far : 0.637 
```
#### Bonus part: Variable importance

We compute the variable importance on the random forest model

```{r}
# Variable importance random forest
rf <- randomForest(share_label ~ ., data = final_df.tr)
varImpPlot(rf)
```

Gini coefficient is an impurity measure. The interpretation of the plot is the following: on the top the most important variables that contribute the most to decrease the node impurity. self_reference.avg_sharess is the most important variable and "day_is_weekend" is the less important.


Variable importance on Random forest


```{r}
# variable importance CART

# without shuffle 
final_df.tree <- rpart(share_label ~ ., data=final_df.tr)
rpart.plot(final_df.tree)
final_df.tree.pred <- predict(final_df.tree, newdata = final_df.te, type = "class")
confusionMatrix(data = as.factor(final_df.tree.pred),
                reference = final_df.te$share_label)


#shuffle on self_reference_avg_sharess
set.seed(143)
final_df.te_shuffle <- final_df.te
final_df.te_shuffle$self_reference_avg_sharess <- sample(final_df.te_shuffle$self_reference_avg_sharess)
final_df.tree.pre_shuffle <- predict(final_df.tree, newdata = final_df.te_shuffle, type = "class")
confusionMatrix(data = as.factor(final_df.tree.pre_shuffle), reference = final_df.te$share_label)

```
Looking at the output we see that the accuracy changes from 0.613 to 0.566 ( the latter result is in the shuffle procedure).


### Multi-Classes : Searching for outstanding popularity


```{r}
set.seed(1996) ## for replication purpose
#split the data set in training/test set
c.index.tr <- createDataPartition(y = classes_df$share_classes,
                                  p = 0.8,
                                  list = FALSE)
classes_df.tr <- classes_df[c.index.tr,]
classes_df.te <- classes_df[-c.index.tr,]
```

Sub sampling
```{r}
df.tr.unpop <- filter(classes_df.tr, share_classes == "Unpopular")
df.tr.pop <- filter(classes_df.tr, share_classes == "Popular")
df.tr.out <- filter(classes_df.tr, share_classes == "Outstanding")
n.out <- min(table(classes_df.tr$share_classes)) #1582
index.unpop <- sample(size=n.out, x=1:nrow(df.tr.unpop), replace=FALSE) ## sub-sample 840 instances from the "No"
index.pop <- sample(size=n.out, x=1:nrow(df.tr.pop), replace=FALSE) ## sub-sample 840 instances from the "No"
df.tr.subs <- data.frame(rbind(df.tr.out,
                               df.tr.unpop[index.unpop, ],
                               df.tr.pop[index.pop, ]))
table(df.tr.subs$share_classes)
```

##### Random forest
In the following part, we computed a random forest on multi-class variable.

```{r}
rf_outstanding <- train(share_classes~.,
            data = df.tr.subs,
            method = "rf",
            metric = "Accuracy",
            tuneGrid = expand.grid(.mtry = 2.5),
            #use the hyperparameter that we've find in
            #the hyperparameter tuning (chunk code line 153-195)
            trControl = trControl)

rf.te.pred <- predict(rf_outstanding,
                      newdata = classes_df.te,
                      type = "raw")

confusionMatrix(data = as.factor(rf.te.pred),
                reference = classes_df.te$share_classes)

```
From the confusion matrix we notice that the overall accuracy is really low. ( 0.482) 
Looking at the results for the different classes: the most predictable papers are the ones with an high level of popularity ( outstanding ). Outstanding class predictions gives us an accuracy of 0.65.

prediction using type = "class" in prediction function
```{r}
library(randomForest)
rf <- randomForest(share_classes ~ ., data = df.tr.subs)
rf.te.pred <- predict(rf, newdata = classes_df.te, type = "class")
confusionMatrix(data = as.factor(rf.te.pred),
                reference = classes_df.te$share_classes)
```
results are slightly worse than the case type = raw.

