# Section 3 : EDA
```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```

<br>

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

We first observe dummy variables that we transformed into categorical levels. This allows us to know what's the type of content that appears the most often. We produce a barplot to represent this and we see that the number of articles written about the world and technology is higher than the others. We can also see that articles about social media and lifestyle are not as frequent. 

</div>

<br>

```{r}
normal_df %>%
  pivot_longer(
    data_channel_is_lifestyle:data_channel_is_world, #Pivot the dataset 
    names_to = "channel",
    values_to = "isnot"
  ) %>%
  filter(isnot == 1) %>% #Filter for values equal to 1
  ggplot(aes(x = channel)) + #Create a barplot 
  geom_bar() +
  coord_flip()
```
<br>

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

In the first look that we had at the data, we saw that our "shares" variable was right-skewed, so we decided to log transform it, so we can have a more "normalized" variable. Here we can see the distribution under the log transformation:

</div>

```{r}
#log_transformation
df$shares <- log(df$shares)
hist(
  df$shares,
  freq = FALSE,
  col = "grey",
  main = "Distribution of Log transformed Shares",
  xlab = "Log_shares"
) #histogram
```
<br>

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

As said in the last section, we transformed our "shares" variable into a categorical one with two levels. The two levels we created, unpopular (<quantile 0.5) and popular(>quantile 0.5) represent 100% of the shares, and in this graphical representation, we see that we mostly have even numbers for these levels, but with a slight advantage for the popular one.

</div>

```{r}
#by label
final_df %>%
  group_by(share_label) %>%
  summarise(n = n()) %>% ungroup() %>% 
  mutate(share_label = reorder(share_label, n)) %>% #Create new column share_label
  ggplot(aes(x = share_label, y = n)) +
  geom_col() +
  geom_segment(aes( #Assign aestethitcs 
    x = share_label ,
    xend = share_label,
    y = 0,
    yend = n
  ), color = "grey") +
  geom_point(size = 3, color = "#69b3a2") +
  coord_flip() +
  theme(
    panel.grid.minor.y = element_blank(),
    panel.grid.major.y = element_blank(),
    legend.position = "none"
  ) +
  xlab("")
```
<br>

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

We then take a look at a certain number of variable distributions given the popularity levels. We see that for the variables "n_tokens_title","n_unique_tokens","num_keywords","n_tokens_content" it is almost a 50-50 split between the unpopular and popular articles. This can indicate that the number of words in the content or the title isn't a good indicator to know if an article is going to be shared a lot or not. On the other hand, the variables "num_hrefs","num_imgs","num_videos","self_reference_avg_sharess" are higher for the popular articles. This can lead us to think that adding links, images, and videos to our article is going to help increase shares. 

We also look at the sentiment analysis of these articles, and we see that for the popular articles, there is a higher polarity in the title, which means that tthey are more positive and that people are going to be keener to share these types of articles. Regarding subjectivity, we see that there isn't much difference between the popular and unpopular articles and that the articles tend to be more objective. 

</div>


```{r}
par(mfrow = c(5, 2))

p1 <- final_df %>%
  group_by(share_label) %>%
  summarise(mean(n_tokens_title)) %>%
  ggplot(aes(x = share_label, y = `mean(n_tokens_title)`)) +
  geom_bar(stat = "identity") +
  ggtitle("") +
  xlab("Popularity Label") +
  ylab("num. token title") +
  theme(plot.title = element_text(hjust = 0.5)) + coord_flip()


p2 <- final_df %>%
  group_by(share_label) %>%
  summarise(mean(n_tokens_content)) %>%
  ggplot(aes(x = share_label, y = `mean(n_tokens_content)`)) +
  geom_bar(stat = "identity") +
  ggtitle("") +
  xlab("Popularity Label") +
  ylab("num. token article") +
  theme(plot.title = element_text(hjust = 0.5)) + coord_flip()#BARPLOT 

p3 <- final_df %>%
  group_by(share_label) %>%
  summarise(mean(n_unique_tokens)) %>%
  ggplot(aes(x = share_label, y = `mean(n_unique_tokens)`)) +
  geom_bar(stat = "identity") +
  ggtitle("") +
  xlab("Popularity Label") +
  ylab("Token rate") +
  theme(plot.title = element_text(hjust = 0.5)) + coord_flip()#BARPLOT 

p4 <- final_df %>%
  group_by(share_label) %>%
  summarise(mean(num_hrefs)) %>% 
  ggplot(aes(x = share_label, y = `mean(num_hrefs)`)) +
  geom_bar(stat = "identity") +
  ggtitle("") +
  ylab("Links") +
  theme(plot.title = element_text(hjust = 0.5)) + coord_flip() #BARPLOT 



p5 <- final_df %>%
  group_by(share_label) %>%
  summarise(mean(num_imgs)) %>%
  ggplot(aes(x = share_label, y = `mean(num_imgs)`)) +
  geom_bar(stat = "identity") +
  ggtitle("") +
  ylab("Images") +
  theme(plot.title = element_text(hjust = 0.5)) + coord_flip()#BARPLOT 



p6 <- final_df %>%
  group_by(share_label) %>%
  summarise(mean(num_videos)) %>%
  ggplot(aes(x = share_label, y = `mean(num_videos)`)) +
  geom_bar(stat = "identity") +
  ggtitle("") +
  ylab("Videos") +
  theme(plot.title = element_text(hjust = 0.5)) + coord_flip()#BARPLOT 



p7 <- final_df %>%
  group_by(share_label) %>%
  summarise(mean(num_keywords)) %>%
  ggplot(aes(x = share_label, y = `mean(num_keywords)`)) +
  geom_bar(stat = "identity") +
  ggtitle("") +
  ylab("Keyword") +
  theme(plot.title = element_text(hjust = 0.5)) + coord_flip() #BARPLOT 



p8 <- final_df %>%
  group_by(share_label) %>%
  summarise(mean(self_reference_avg_sharess)) %>%
  ggplot(aes(x = share_label, y = `mean(self_reference_avg_sharess)`)) +
  geom_bar(stat = "identity") +
  ggtitle("") +
  ylab("Reference") +
  theme(plot.title = element_text(hjust = 0.5)) + coord_flip() #BARPLOT 


#Sentiment analysis is basically the process of determining the attitude or the emotion of the writer, i.e., whether it is positive or negative or neutral.
#The sentiment function of textblob returns two properties, polarity, and subjectivity.
#Polarity is float which lies in the range of [-1,1] where 1 means positive statement and -1 means a negative statement.
#Subjective sentences generally refer to personal opinion, emotion or judgment whereas objective refers to factual information.
#Subjectivity is also a float which lies in the range of [0,1].

#polarity and subjectivity given shares popularity classes

p9 <- final_df %>% group_by(share_label) %>% summarise(mean(title_sentiment_polarity)) %>%
  ggplot(aes(x = share_label, y = `mean(title_sentiment_polarity)`)) +
  geom_bar(stat = "identity") +
  ylab("Polarity") +
  theme(plot.title = element_text(hjust = 0.5)) + coord_flip()#BARPLOT 

p10 <- final_df %>%
  group_by(share_label) %>%
  summarise(mean(title_subjectivity)) %>%
  ggplot(aes(x = share_label, y = `mean(title_subjectivity)`)) +
  geom_bar(stat = "identity") +
  ylab("Subjectivity") +
  theme(plot.title = element_text(hjust = 0.5)) + coord_flip()#BARPLOT 

p1 + p2 + p3 + p4 + p5 + p6 + p7 + p8 + p9 + p10
```
<br>

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">


Lastly, we used boxplots to see if there is a difference among the categorical variables for the number of shares, but as we can see, the number of shares isn't really affected by the type of content,the day the article is published or by the fact that it's published on a weekend.

</div>


```{r}
par(mfrow = c(3, 1))
b1 <- ggplot(data = new, aes(data_channel, log(shares))) + geom_boxplot() + coord_flip() #Create a boxplot of channels given the shares

b2 <- ggplot(data = new, aes(day, log(shares))) + geom_boxplot()+ coord_flip() #Create a boxplot of days given the shares 

b3 <- ggplot(data = new, aes(day_is_weekend, log(shares))) + geom_boxplot()+ coord_flip() ##Create a boxplot of weekend given the shares

b1 + b2 + b3
```
